---
title: "cmdstanr"
output:
  html_document:
    df_print: paged
---

This is an example on how to use [cmdstanr](https://mc-stan.org/cmdstanr/).

We will run it on CPU as well as GPU to compare the performance 

# Our model

We will use `bernoulli_logit_glm` (model from the `cmdstanr` vignette)[https://raw.githubusercontent.com/stan-dev/cmdstanr/master/vignettes/articles-online-only/opencl-files/bernoulli_logit_glm.stan] and have saved it in this folder as  `model.stan`. 

# Data generation

```{r}
n <- 250000
k <- 20
X <- matrix(rnorm(n * k), ncol = k)
y <- rbinom(n, size = 1, prob = plogis(3 * X[,1] - 2 * X[,2] + 1))
mdata <- list(k = k, n = n, y = y, X = X)
```


# CPU run

## Prerequisites (one-time setup)

First of all, you need to install the `cmdstanr` package. This can be achieved by running 

```{r message=FALSE, warning=FALSE}
renv::install('cmdstanr', repos = c('https://stan-dev.r-universe.dev', getOption("repos")), prompt=FALSE)
```

Then you need to install `cmdstan`. We are making use of all the cores we have available via the `parallelly` package. Compiling C++ code is inherently slow so let's make sure we use all available resources. Additionally we are specifying a specific `cmdstan` version for both good practice and increased reproducibility. 

```{r message=FALSE, warning=FALSE}
library(cmdstanr)
if (!dir.exists(paste0(Sys.getenv("HOME"),"/.cmdstan/cmdstan-2.34.1"))) {
cmdstanr::install_cmdstan(
  version = "2.34.1", 
  cores = parallelly::availableCores()
  )
}
```

Please note you always can rebuild cmdstan via the `rebuild_cmdstan()` function. 


## Compile the model 

```{r}
cmdstanr::set_cmdstan_path(paste(Sys.getenv("HOME"),".cmdstan/cmdstan-2.34.1",sep="/"))
mod_cpu <- cmdstanr::cmdstan_model("model.stan",compile=FALSE)
mod_cpu$exe_file("model-cpu")
mod_cpu$compile()
```

## Run the model

```{r}
time_cpu<-system.time(fit_cpu <- mod_cpu$sample(data = mdata, chains = 4, parallel_chains = 4, refresh = 0))
```

# GPU run 

## Prerequisites (one-time setup)

`cmdstan` uses OpenCL to make use of a GPU. We will need to build `cmdstan` and link against OpenCL. 

```{r}
cmdstan_dir = "~/.cmdstan-gpu"

if (!dir.exists(paste0(cmdstan_dir, "/cmdstan-2.34.1"))) {
  cpp_options = list("LDFLAGS+= -lOpenCL")
  
  
  dir.create(cmdstan_dir)
  cmdstanr::install_cmdstan(
    version = "2.34.1",
    dir = cmdstan_dir,
    cpp_options = cpp_options,
    cores = parallelly::availableCores()
  )
}
```

For each `cmdstan` execution leveraging GPUs we now have to point to this `cmdstan` install

```{r}
cmdstanr::set_cmdstan_path(paste(cmdstan_dir,"cmdstan-2.34.1",sep="/"))
```

## Compile the model 

```{r}
cmdstanr::set_cmdstan_path(paste(cmdstan_dir,"cmdstan-2.34.1",sep="/"))
mod_gpu <- cmdstanr::cmdstan_model("model.stan",compile=FALSE)
mod_gpu$exe_file("model-gpu")
mod_gpu$compile(cpp_options = list(stan_opencl = TRUE))
```

## Run the model

```{r}
time_gpu<-system.time(fit_gpu <- mod_gpu$sample(data = mdata, chains = 4, parallel_chains = 4, refresh = 0))
```

# Compare CPU with GPU

```{r}
time_cpu/time_gpu
```